ms_5

### 2020-04-20 ms_5.Rmd
newdatasets and graphs!!

library(dplyr)
library(janitor)
library(tidyverse)
```

@@

### Read in Data
drugoverdosedeaths<- read.csv("ms_6/VSRR_Provisional_Drug_Overdose_Death_Counts.csv")

@@

        ggplot(drugoverdosedeaths, aes(x= reorder(state_name,-data_value), y= data_value)) +
            geom_col() +
            facet_wrap(~time)
            labs(title = "Drug Overdose Deaths by State Over Time", 
                 x = "State", 
                 y = "Drug Overdose Deaths") +

@@

>>>>

library(dplyr)
library(janitor)
library(tidyverse)
library(ggthemes)
library(maps)
library(leaflet)
```

@@

drugoverdosedeaths<- read.csv("ms_6/VSRR_Provisional_Drug_Overdose_Death_Counts.csv")

@@

drugoverdosedeaths2 <- drugoverdosedeaths %>%
 filter(time == "January 2015")

        ggplot(drugoverdosedeaths2, aes(x= reorder(state_name,-data_value), y= data_value)) +
            geom_col() +
            labs(title = "Drug Overdose Deaths by State Over Time", 
                 x = "State", 
                 y = "Drug Overdose Deaths") +

@@

```{r, message=FALSE}
#drug treatment centers

treatment_locations <- read_csv("Drug Treatment Centers.csv") %>%
  select(`Program Name`, Street, City, State, Zipcode)

us_state <- map_data("state") %>%
  rename(state_full = region)

# https://simplemaps.com/data/us-cities

longlatinfo <- read_csv("uscitieslonglat.csv") %>%
  rename(City = city) %>%
  rename(State = state_id)

treatment_locations_map <- treatment_locations %>%
  left_join(longlatinfo, by = c("City", "State")) %>%
  group_by(State) %>%
  mutate(count = n()) %>%
  mutate(state_name = tolower(state_name)) %>%
  rename(state_full = state_name) %>%
  left_join(us_state, by = "state_full")

ggplot(data = treatment_locations_map,
            mapping = aes(x = long, y = lat.y,
                          fill = count, group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,20,40,60,80,120)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom") +
  labs(title = "Distribution of Opioid Treatment Centers in 2020",
       color = "Number of Treatment Centers in State")

# per capita pop data taken from https://www.census.gov/data/tables/time-series/demo/popest/2010s-state-total.html#par_textimage

state_pop <- read_csv("State Population Estimate.csv") %>%
  select(NAME, POPESTIMATE2019) %>%
  rename(state_full = NAME) %>%
  mutate(state_full = tolower(state_full))

treatment_locations_map_per_cap <- treatment_locations_map %>%
  left_join(state_pop, by = "state_full") %>%
  group_by(State) %>%
  mutate(count_per_pop = n()/POPESTIMATE2019)

ggplot(data = treatment_locations_map_per_cap,
            mapping = aes(x = long, y = lat.y,
                          fill = count_per_pop, group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,20,40,60,80,120)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom") +
  labs(title = "Distribution of Opioid Treatment Centers in 2020 (Per Capita)",
       color = "Number of Treatment Centers in State Per Capita")
```

```{r, message = FALSE}
#MA specific data

madeathbyagegender <- read_csv("MAFatalOpioidOverdosesByAgeAndGender.csv")

macountyheroin <- read_csv("MAHeroinAsPrimarySubstanceOfUse.csv")

madeathbyrace <- read_csv("MAOpioidDeathsByRaceAndEthnicity.csv")

matypeofsubstance <- read_csv("MAPrimarySubstanceofUseWhenEnteringTreatmentbyTown.csv")

toxdeath <- read_csv("MAProportionDecedentsWithRxHistorybyToxicologyScreen.csv",
         skip = 1)

```{r}
mavsus_death <- read_csv("MAAgeAdjustedOpioidRelatedDeathRateByYear.csv")

mavsus_death <- mavsus_death %>%
  group_by(Geography) %>%
  filter (Year != "2015") %>%
  filter (Year != "1999") %>%
  rename(deathperhundredthousand =`Age-Adjusted Opioid-Related Death Rate per 100,000 People`)

ggplot(mavsus_death, aes(x = Year, y = as.numeric(deathperhundredthousand), color = Geography)) +
  geom_line() +
theme_classic() +
  labs(x = "Year",
       y = "Deaths per 100,000",
       title = "Comparison of Death Rates between Massachusetts and the United States",
       subtitle = "Between the Years of 2000 and 2014") 
```

```{r}
counties <- read_csv("zip_codes_states.csv") %>%
  filter(state == "MA") %>%
  rename(County = county) %>%
  rename(Municipality = city)

madeathbycounty <- read_csv("MAAverageAnnualOpioidRelatedDeathRateper100,000People.csv")

countypop <- read_csv("countypop.csv") %>%
  mutate(Pop = Pop/100000) %>%
  rename(subregion = CTYNAME) %>%
  mutate(subregion = tolower(subregion)) %>%
  select(subregion, Pop)

madeathbycountywlonglat <- madeathbycounty %>%
  left_join(counties, by = "Municipality")

madeathbycountywlonglat <- madeathbycountywlonglat %>%
  select(County,
         `Confirmed Opioid Related Death Count 2001-2005`,
         `Confirmed Opioid Related Death Count 2006-2010`,
         `Confirmed Opioid Related Death Count 2011-2015`,
         `latitude`,
         `longitude`,
         `Municipality`) %>%
  distinct(Municipality, .keep_all= TRUE) %>%
  na.omit() %>%
  rename(subregion = County) %>%
  mutate(subregion = tolower(subregion)) %>%
  group_by(subregion) %>%
  mutate(total_deaths_2001.5 = sum(`Confirmed Opioid Related Death Count 2001-2005`)) %>%
  mutate(total_deaths_2006.10 = sum(`Confirmed Opioid Related Death Count 2006-2010`)) %>%
  mutate(total_deaths_2011.15 = sum(`Confirmed Opioid Related Death Count 2011-2015`)) %>%
  distinct(subregion, .keep_all= TRUE)

us_county <- map_data("county") %>%
  filter(region == "massachusetts")

# https://www.indexmundi.com/facts/united-states/quick-facts/massachusetts/percent-of-people-of-all-ages-in-poverty#table

povertybycounty <- read_csv("PovertyByCounty.csv") %>%
  rename(subregion = County) %>%
  mutate(subregion = tolower(subregion))

full_data <- us_county %>%
  left_join(madeathbycountywlonglat, by = "subregion") %>%
  left_join(countypop, by = "subregion") %>%
  left_join(povertybycounty, by = "subregion") %>%
  mutate(percap_2001.5 = total_deaths_2001.5/Pop) %>%
  mutate(percap_2006.10 = total_deaths_2006.10/Pop) %>%
  mutate(percap_2011.15 = total_deaths_2011.15/Pop)

ggplot(full_data, aes(x = `Poverty Percent`, y = percap_2001.5)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

full_data %>%
  summarize(correlation = cor(percap_2001.5,`Poverty Percent`))

ggplot(full_data, aes(x = `Poverty Percent`, y = percap_2006.10)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

full_data %>%
  summarize(correlation = cor(percap_2006.10,`Poverty Percent`))

ggplot(full_data, aes(x = `Poverty Percent`, y = percap_2011.15)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

full_data %>%
  summarize(correlation = cor(percap_2011.15,`Poverty Percent`))

ggplot(data = full_data,
            mapping = aes(x = long, y = lat,
                          fill = `Poverty Percent`,
                          group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,2,4,6,8,10)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")

ggplot(data = full_data,
            mapping = aes(x = long, y = lat,
                          fill = percap_2001.5,
                          group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,20,40,60,80,100)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")

ggplot(data = full_data,
            mapping = aes(x = long, y = lat,
                          fill = percap_2006.10,
                          group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,20,40,60,80,100)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")

ggplot(data = full_data,
            mapping = aes(x = long, y = lat,
                          fill = percap_2011.15,
                          group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,20,40,60,80,100)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")

# don't put these in final app

ggplot(data = full_data,
            mapping = aes(x = long, y = lat,
                          fill = total_deaths_2001.5,
                          group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,100,200,300,400,500,600)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")

ggplot(data = full_data,
            mapping = aes(x = long, y = lat,
                          fill = total_deaths_2006.10,
                          group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,100,200,300,400,500,600)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")

ggplot(data = full_data,
            mapping = aes(x = long, y = lat,
                          fill = total_deaths_2011.15,
                          group = group)) + 
  geom_polygon(color = "gray90", size = 0.05) + 
  theme_map() +
  scale_fill_gradient(low = "white", high = "#CB454A",
                      breaks = c(0,100,200,300,400,500,600)) +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")

#treatment centers in MA

treatment_locations_map_mas <- treatment_locations %>%
  left_join(longlatinfo, by = c("City", "State")) %>%
  filter(State == "MA") %>%
  drop_na()

leaflet(options = leafletOptions(dragging = TRUE,
                                 minZoom = 8, 
                                 maxZoom = 10)) %>%
  addProviderTiles("CartoDB") %>%
  addCircleMarkers(data = treatment_locations_map_mas,
                   radius = 3,
                   label = ~`Program Name`)

# need to figure this out bc color isnt working
 # addLegend(position = "bottomright", colors = "#0000FF", title = "Opioid Treatment Sites")
